---
title: "Data Visualization in R"
author: "Jenn Schilling"
date: "3/2/2021"
output: html_document
---

```{r setup, include=FALSE}

# Set defualt chunk options
knitr::opts_chunk$set(echo = TRUE, message = FALSE) 

# Set the default size of figures
# knitr::opts_chunk$set(fig.width = 8, fig.height = 5)  

library(here)
library(tidyverse)

```


Get IPEDS directory information and select a group of institutions.

```{r get-inst-ids}

directory <- read_csv(here('data', 'ipeds_2019', 'hd2019.csv'))

inst_list <- directory %>%
  janitor::clean_names() %>%
  filter(instnm %in% c(
    'University of Arizona',
    'University of California-Davis',
    'University of California-Los Angeles',
    'University of Florida',
    'University of Illinois at Urbana-Champaign',
    'University of Iowa',
    'University of Maryland-College Park',
    'Michigan State University',
    'University of Minnesota-Twin Cities',
    'University of North Carolina at Chapel Hill',
    'Ohio State University-Main Campus',
    'Pennsylvania State University-Main Campus',
    'Texas A & M University-College Station',
    'The University of Texas at Austin',
    'University of Washington-Seattle Campus',
    'University of Wisconsin-Madison')) %>% 
  mutate(control = "Public", # All the same
         hdegrofr1 = "Doctor's degree - research/scholarship and professional practice", # All the same
         hospital = ifelse(hospital == 1, "Yes", "No"),
         medical = ifelse(medical == 1, "Yes", "No"),
         locale = case_when(
           locale == 11 ~ "City: Large",
           locale == 12 ~ "City: Midsize",
           locale == 13 ~ "City: Small",
           locale == 21 ~ "Suburb: Large",
           locale == 23 ~ "Suburb: Small",
           TRUE ~ "Other" ),
         instsize = "20,000 and above", # All the same
         landgrnt = ifelse(landgrnt == 1, "Land Grant Institution", "Not Land Grant Institution"),
         c18basic = "Doctoral Universities: Very High Research Activity", # All the same
         c18enprf = case_when(
           c18enprf == 4 ~ "High undergraduate",
           c18enprf == 5 ~ "Majority undergraduate",
           TRUE ~ "Other"),
         c18szset = case_when(
           c18szset == 15 ~ "Four-year, large, primarily nonresidential",
           c18szset == 16 ~ "Four-year, large, primarily residential",
           c18szset == 17 ~ "Four-year, large, highly residential",
           TRUE ~ "Other")
         ) %>%
  select(unitid, instnm, city, stabbr, control, hdegrofr1, hospital, medical, locale, instsize, landgrnt, c18basic, c18enprf, c18szset)

```


Get the IPEDS Admissions data for institutions of interest for survey years 2017 - 2019.

```{r get-adm-data}

adm_17 <- read_csv(here('data', 'ipeds_2017', 'adm2017_rv.csv')) %>% janitor::clean_names() %>% mutate(year = 2017)
adm_18 <- read_csv(here('data', 'ipeds_2018', 'adm2018_rv.csv')) %>% janitor::clean_names() %>% mutate(year = 2018)
adm_19 <- read_csv(here('data', 'ipeds_2019', 'adm2019.csv')) %>% janitor::clean_names() %>% mutate(year = 2019)

adm <- rbind(adm_17, adm_18, adm_19)

inst_adm <- left_join(inst_list, adm, by = 'unitid')

# Process Admissions Data
inst_adm <- inst_adm %>%
  rename(n_applicants = applcn,
         n_admits = admssn,
         n_enrl = enrlt,
         n_sat = satnum,
         pct_sat = satpct,
         n_act = actnum,
         pct_act = actpct,
         test_req = admcon7) %>%
  mutate(sat_score = (satvr25 + satvr75) / 2 + (satmt25 + satmt75) / 2,
         act_score = actcm25 + actcm75 / 2,
         test_req = case_when(
           test_req == 1 ~ "Required",
           test_req == 5 ~ "Considered by not required",
           test_req == 2 ~ "Recommended",
           TRUE ~ "Other"
         ),
         year = as.integer(year)) %>%
  select(year, unitid, n_applicants, n_admits, n_enrl, test_req, n_sat, pct_sat, sat_score, n_act, pct_act, act_score)

```


Get the IPEDS Enrollment data for institutions of interest for survey years 2017 - 2019.

```{r get-enrlmnt-data}

enrlmnt_17 <- read_csv(here('data', 'ipeds_2017', 'ef2017a_rv.csv')) %>% janitor::clean_names() %>% mutate(year = 2017)
enrlmnt_18 <- read_csv(here('data', 'ipeds_2018', 'ef2018a_rv.csv')) %>% janitor::clean_names() %>% mutate(year = 2018)
enrlmnt_19 <- read_csv(here('data', 'ipeds_2019', 'ef2019a.csv')) %>% janitor::clean_names() %>% mutate(year = 2019)

enrlmnt <- rbind(enrlmnt_17, enrlmnt_18, enrlmnt_19)

inst_enrlmnt <- left_join(inst_list, enrlmnt, by = 'unitid')

# Process Enrollment Data
inst_enrlmnt <- inst_enrlmnt %>%
  filter(efalevel == 1 | # All students total
         efalevel == 2 | # All students undergraduate total
         efalevel == 12 # All students graduate total
           ) %>%
  mutate(lstudy = case_when(
      lstudy == 1 ~ "Undergraduate",
      lstudy == 3 ~ "Graduate",
      lstudy == 4 ~ "All students",
      TRUE ~ "Other")) %>%
  rename(n_enrl = eftotlt,
         n_enrl_m = eftotlm,
         n_enrl_w = eftotlw) %>%
  select(year, unitid, lstudy, n_enrl, n_enrl_m, n_enrl_w) %>%
  pivot_longer(cols = n_enrl:n_enrl_w,
               names_to = "gender",
               values_to = "n_enrl") %>%
  mutate(gender = case_when(
    gender == "n_enrl" ~ "Total",
    gender == "n_enrl_m" ~ "Men",
    gender == "n_enrl_w" ~ "Women"
  ),
  year = as.integer(year))

```


Get the IPEDS Completions data for institutions of interest for survey years 2017 - 2019.

```{r get-compl-data}

compl_17 <- read_csv(here('data', 'ipeds_2017', 'c2017_c_rv.csv')) %>% janitor::clean_names() %>% mutate(year = 2017)
compl_18 <- read_csv(here('data', 'ipeds_2018', 'c2018_c_rv.csv')) %>% janitor::clean_names() %>% mutate(year = 2018)
compl_19 <- read_csv(here('data', 'ipeds_2019', 'c2019_c.csv')) %>% janitor::clean_names() %>% mutate(year = 2019)

compl <- rbind(compl_17, compl_18, compl_19)

inst_compl <- left_join(inst_list, compl, by = 'unitid')

# Process completions data
inst_compl <- inst_compl %>%
  mutate(awlevelc = case_when(
    awlevelc == "05" ~ "Bachelor's degree",
    awlevelc == "07" ~ "Master's degree",
    awlevelc == "09" ~ "Doctor's degree",
    TRUE ~ "Other"
  )) %>%
  filter(awlevelc != "Other") %>%
  rename(n_deg = cstotlt,
         n_deg_m = cstotlm,
         n_deg_w = cstotlw) %>%
  select(year, unitid, awlevelc, n_deg, n_deg_m, n_deg_w) %>%
  pivot_longer(cols = n_deg:n_deg_w,
               names_to = "gender",
               values_to = "n_deg") %>%
  mutate(gender = case_when(
    gender == "n_deg" ~ "Total",
    gender == "n_deg_m" ~ "Men",
    gender == "n_deg_w" ~ "Women"
  ),
  year = as.integer(year))

```


**Session 1**
Why is data visualization important? (5 minutes)
Best practices in data visualization (5 minutes)
Data visualization in R (examples of R data visualizations, concept of the grammar of graphics) (5 minutes)
R set-up (Rmarkdown, packages needed, importing data) (10 minutes)
How to make a scatterplot in R (guided exercise) (10 minutes)
How to make a line graph in R (guided exercise) (10 minutes)
Questions (15 minutes)

**Session 2**
Recap of previous session (grammar of graphics, R set-up, making a plot) (5 minutes)
How to make a bar graph in R (guided exercise) (10 minutes)
Adding and changing titles and labels in R (10 minutes)
Formatting options for plots (including how to make small multiples in R) (10 minutes)
How to save a plot in R (5 minutes)
Where to go from here (5 minutes)
Questions (15 minutes)


```{r scatter-plot}

inst_adm_plot <- left_join(inst_list, inst_adm, by = "unitid") %>%
  filter(year == 2019)

ggplot(data = inst_adm_plot,
       mapping = aes(x = sat_score,
                     y = act_score)) +
  geom_point()

```

```{r line-graph}

inst_adm_plot <- left_join(inst_list, inst_adm, by = "unitid")

ggplot(data = inst_adm_plot,
       mapping = aes(x = year,
                     y = sat_score)) +
  geom_line()

# Need to add group mapping
ggplot(data = inst_adm_plot,
       mapping = aes(x = year,
                     y = sat_score,
                     group = unitid)) +
  geom_line()

# What about number who submitted by year?
ggplot(data = inst_adm_plot,
       mapping = aes(x = year,
                     y = n_sat,
                     group = unitid)) +
  geom_line()

# Who's that outlier in 2019?
ggplot(data = inst_adm_plot,
       mapping = aes(x = year,
                     y = n_sat,
                     group = unitid)) +
  geom_line() +
  geom_text(mapping = aes(label = instnm)) # Add Text Labels


# We can see it's Penn State, but the labels are kind of a mess
ggplot(data = inst_adm_plot,
       mapping = aes(x = year,
                     y = n_sat,
                     group = unitid)) +
  geom_line() +
  # Filter to only the last year for the text labels
  geom_text(data = inst_adm_plot %>% filter(year == 2019),
            mapping = aes(x = year,
                          y = n_sat,
                          label = instnm))

# How can we make the labels more readable?
ggplot(data = inst_adm_plot,
       mapping = aes(x = year,
                     y = n_sat,
                     group = unitid)) +
  geom_line() +
  # Change the x mapping for the text labels and add some format adjustments 
  geom_text(data = inst_adm_plot %>% filter(year == 2019),
            mapping = aes(x = Inf,
                          y = n_sat,
                          label = instnm),
            hjust = -.01) +
  coord_cartesian(clip = "off") +
  theme(plot.margin = margin(0, 250, 0, 0))


```
```{r bar-plot}

# How many institutions of each type?
ggplot(data = inst_list,
       mapping = aes(x = locale)) +
  geom_bar()

# Can we sort them?
# Yes, by using the fct_infreq() function from the forcats package in tidyverse
# fct_infreq() reorders the levels by their frequencies
ggplot(data = inst_list,
       mapping = aes(x = fct_infreq(locale))) +
  geom_bar()

# What if we want to make a bar plot with something other than counts?

inst_enrlmnt_plot <- left_join(inst_list, inst_enrlmnt, by = "unitid")

# What does enrollment look like by urban area?

# This will give total enrollment in each urban area
ggplot(data = inst_enrlmnt_plot %>% 
         filter(lstudy == "All students", gender == "Total", year == 2019),
       mapping = aes(x = locale,
                     y = n_enrl)) +
  geom_col()

# See - the enrollments for each institution are stacked on top of one another
ggplot(data = inst_enrlmnt_plot %>% 
         filter(lstudy == "All students", gender == "Total", year == 2019),
       mapping = aes(x = locale,
                     y = n_enrl,
                     fill = instnm)) +
  geom_col()

# What if we want the average enrollment in each area? 
inst_enrlmnt_plot_sub <- inst_enrlmnt_plot %>%
  filter(lstudy == "All students", gender == "Total", year == 2019) %>%
  group_by(locale) %>%
  summarize(avg_enrl = mean(n_enrl),
            .groups = "drop")

ggplot(data = inst_enrlmnt_plot_sub,
       mapping = aes(x = locale,
                     y = avg_enrl)) +
  geom_col()

# Sort by average enrollment
# This time, use the reorder() function to arrange the locale column by the avg_enrl column
ggplot(data = inst_enrlmnt_plot_sub,
       mapping = aes(x = reorder(locale, -avg_enrl),
                     y = avg_enrl)) +
  geom_col()

# What if we want to see the enrollment by institution and urban area?
ggplot(data = inst_enrlmnt_plot %>% 
         filter(lstudy == "All students", gender == "Total", year == 2019),
       mapping = aes(x = n_enrl,
                     y = reorder(instnm, n_enrl))) +
  geom_col() +
  facet_wrap(~ locale,
             scales = "free_y",
             ncol = 1)

```

```{r fill-color}

# How has average enrollment changed by urban area over time?
inst_enrlmnt_plot_sub <- inst_enrlmnt_plot %>%
  filter(lstudy == "All students", gender == "Total") %>%
  group_by(year, locale) %>%
  summarize(avg_enrl = mean(n_enrl),
            .groups = "drop")

ggplot(data = inst_enrlmnt_plot_sub,
       mapping = aes(x = year,
                     y = avg_enrl,
                     fill = locale)) +
  geom_col() 


# Remove the stacking
ggplot(data = inst_enrlmnt_plot_sub,
       mapping = aes(x = year,
                     y = avg_enrl,
                     fill = locale)) +
  geom_col(position = "dodge") 

# Make a different plot for each urban area
ggplot(data = inst_enrlmnt_plot_sub,
       mapping = aes(x = year,
                     y = avg_enrl,
                     fill = locale)) +
  geom_col(position = "dodge") +
  facet_wrap(~ locale) +
  guides(fill = FALSE) # remove the legend

# Maybe it would be better to show this with lines?
ggplot(data = inst_enrlmnt_plot_sub,
       mapping = aes(x = year,
                     y = avg_enrl,
                     fill = locale)) +
  geom_line() +
  facet_wrap(~ locale) +
  guides(fill = FALSE) 

# What happened to the color?
# Bars, histograms, and boxplots use fill; lines and points use color
ggplot(data = inst_enrlmnt_plot_sub,
       mapping = aes(x = year,
                     y = avg_enrl,
                     color = locale)) +
  geom_line() +
  facet_wrap(~ locale) +
  guides(color = FALSE) 

```


```{r formatting-1}

inst_compl_plot <- left_join(inst_list, inst_compl, by = "unitid") %>%
  mutate(city_sub = str_extract(locale, ".+?(?=:)"))

ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total", year == 2019),
       mapping = aes(x = n_deg,
                     y = instnm,
                     fill = city_sub)) +
  geom_col() +
  facet_wrap(~ awlevelc)

# Reorder degree levels
ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total", year == 2019),
       mapping = aes(x = n_deg,
                     y = instnm,
                     fill = city_sub)) +
  geom_col() +
  facet_wrap(~ factor(awlevelc,
                      levels = c("Bachelor's degree",
                                 "Master's degree",
                                 "Doctor's degree")))

# Reorder institutions by city/suburb 
ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total", year == 2019),
       mapping = aes(x = n_deg,
                     y = fct_reorder(instnm, city_sub),
                     fill = city_sub)) +
  geom_col() +
  facet_wrap(~ factor(awlevelc,
                      levels = c("Bachelor's degree",
                                 "Master's degree",
                                 "Doctor's degree")))

# Add titles and labels
ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total", year == 2019),
       mapping = aes(x = n_deg,
                     y = fct_reorder(instnm, city_sub),
                     fill = city_sub)) +
  geom_col() +
  facet_wrap(~ factor(awlevelc,
                      levels = c("Bachelor's degree",
                                 "Master's degree",
                                 "Doctor's degree"))) +
  labs(title = "Number of Degrees Awarded by Institution",
       subtitle = "Academic Year 2018-2019",
       x = "",
       y = "")

# Move legend to the bottom and remove legend title
ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total", year == 2019),
       mapping = aes(x = n_deg,
                     y = fct_reorder(instnm, city_sub),
                     fill = city_sub)) +
  geom_col() +
  facet_wrap(~ factor(awlevelc,
                      levels = c("Bachelor's degree",
                                 "Master's degree",
                                 "Doctor's degree"))) +
  labs(title = "Number of Degrees Awarded by Institution",
       subtitle = "Academic Year 2018-2019",
       x = "",
       y = "",
       fill = "") +
  theme(legend.position = "bottom")

# Add note about data source
ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total", year == 2019),
       mapping = aes(x = n_deg,
                     y = fct_reorder(instnm, city_sub),
                     fill = city_sub)) +
  geom_col() +
  facet_wrap(~ factor(awlevelc,
                      levels = c("Bachelor's degree",
                                 "Master's degree",
                                 "Doctor's degree"))) +
  labs(title = "Number of Degrees Awarded by Institution",
       subtitle = "Academic Year 2018-2019",
       caption = "Data from IPEDS",
       x = "",
       y = "",
       fill = "") +
  theme(legend.position = "bottom")



```

```{r formatting-2}


# Change the plot theme from the default
ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total", year == 2019),
       mapping = aes(x = n_deg,
                     y = fct_reorder(instnm, city_sub),
                     fill = city_sub)) +
  geom_col() +
  facet_wrap(~ factor(awlevelc,
                      levels = c("Bachelor's degree",
                                 "Master's degree",
                                 "Doctor's degree"))) +
  labs(title = "Number of Degrees Awarded by Institution",
       subtitle = "Academic Year 2018-2019",
       caption = "Data from IPEDS",
       x = "",
       y = "",
       fill = "") +
  theme_minimal() +
  theme(legend.position = "bottom") 

# Add commas to the x-axis labels
ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total", year == 2019),
       mapping = aes(x = n_deg,
                     y = fct_reorder(instnm, city_sub),
                     fill = city_sub)) +
  geom_col() +
  facet_wrap(~ factor(awlevelc,
                      levels = c("Bachelor's degree",
                                 "Master's degree",
                                 "Doctor's degree"))) +
  scale_x_continuous(labels = scales::comma) +
  labs(title = "Number of Degrees Awarded by Institution",
       subtitle = "Academic Year 2018-2019",
       caption = "Data from IPEDS",
       x = "",
       y = "",
       fill = "") +
  theme_minimal() +
  theme(legend.position = "bottom") 


# Change the bar fill colors (using ColorBrewer)
ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total", year == 2019),
       mapping = aes(x = n_deg,
                     y = fct_reorder(instnm, city_sub),
                     fill = city_sub)) +
  geom_col() +
  facet_wrap(~ factor(awlevelc,
                      levels = c("Bachelor's degree",
                                 "Master's degree",
                                 "Doctor's degree"))) +
  scale_x_continuous(labels = scales::comma) +
  scale_fill_brewer(palette = "Dark2") +
  labs(title = "Number of Degrees Awarded by Institution",
       subtitle = "Academic Year 2018-2019",
       caption = "Data from IPEDS",
       x = "",
       y = "",
       fill = "") +
  theme_minimal() +
  theme(legend.position = "bottom") 

# Change the bar fill colors manually
ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total", year == 2019),
       mapping = aes(x = n_deg,
                     y = fct_reorder(instnm, city_sub),
                     fill = city_sub)) +
  geom_col() +
  facet_wrap(~ factor(awlevelc,
                      levels = c("Bachelor's degree",
                                 "Master's degree",
                                 "Doctor's degree"))) +
  scale_x_continuous(labels = scales::comma) +
  scale_fill_manual(values = c("#00A4CCFF", "#F95700FF")) +
  labs(title = "Number of Degrees Awarded by Institution",
       subtitle = "Academic Year 2018-2019",
       caption = "Data from IPEDS",
       x = "",
       y = "",
       fill = "") +
  theme_minimal() +
  theme(legend.position = "bottom") 




```
```{r alternate-formatting-1}

inst_compl_plot <- left_join(inst_list, inst_compl, by = "unitid") 

ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total", unitid == 104179),
       mapping = aes(x = year,
                     y = n_deg,
                     color = awlevelc)) +
  geom_line()

# Reorder degree levels
ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total", unitid == 104179),
       mapping = aes(x = year,
                     y = n_deg,
                     color = factor(awlevelc,
                      levels = c("Bachelor's degree",
                                 "Master's degree",
                                 "Doctor's degree")))) +
  geom_line()

# Add titles and labels
ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total", unitid == 104179),
       mapping = aes(x = year,
                     y = n_deg,
                     color = factor(awlevelc,
                      levels = c("Bachelor's degree",
                                 "Master's degree",
                                 "Doctor's degree")))) +
  geom_line() +
  labs(title = "Number of Degrees Awarded by the University of Arizona by Academic Year",
       x = "Academic Year",
       y = "")

# Move legend to the top and remove legend title
ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total", unitid == 104179),
       mapping = aes(x = year,
                     y = n_deg,
                     color = factor(awlevelc,
                      levels = c("Bachelor's degree",
                                 "Master's degree",
                                 "Doctor's degree")))) +
  geom_line() +
  labs(title = "Number of Degrees Awarded by the University of Arizona by Academic Year",
       x = "Academic Year",
       y = "",
       color = "") +
  theme(legend.position = "top")

# Add note about data source
ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total", unitid == 104179),
       mapping = aes(x = year,
                     y = n_deg,
                     color = factor(awlevelc,
                      levels = c("Bachelor's degree",
                                 "Master's degree",
                                 "Doctor's degree")))) +
  geom_line() +
  labs(title = "Number of Degrees Awarded by the University of Arizona by Academic Year",
       x = "Academic Year",
       y = "",
       color = "",
       caption = "Data from IPEDS Completions Survey") +
  theme(legend.position = "top")


# Add Theme
ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total", unitid == 104179),
       mapping = aes(x = year,
                     y = n_deg,
                     color = factor(awlevelc,
                      levels = c("Bachelor's degree",
                                 "Master's degree",
                                 "Doctor's degree")))) +
  geom_line() +
  labs(title = "Number of Degrees Awarded by the University of Arizona by Academic Year",
       x = "Academic Year",
       y = "",
       color = "",
       caption = "Data from IPEDS Completions Survey") +
  theme_minimal() +
  theme(legend.position = "top")


# Add commas to y-axis labels
ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total", unitid == 104179),
       mapping = aes(x = year,
                     y = n_deg,
                     color = factor(awlevelc,
                      levels = c("Bachelor's degree",
                                 "Master's degree",
                                 "Doctor's degree")))) +
  geom_line() +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "Number of Degrees Awarded by the University of Arizona by Academic Year",
       x = "Academic Year",
       y = "",
       color = "",
       caption = "Data from IPEDS Completions Survey") +
  theme_minimal() +
  theme(legend.position = "top")

# Remove decimals from x-axis labels and make them academic year ranges
ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total", unitid == 104179),
       mapping = aes(x = year,
                     y = n_deg,
                     color = factor(awlevelc,
                      levels = c("Bachelor's degree",
                                 "Master's degree",
                                 "Doctor's degree")))) +
  geom_line() +
  scale_y_continuous(labels = scales::comma) +
  scale_x_continuous(breaks = c(2017, 2018, 2019)) +
  labs(title = "Number of Degrees Awarded by the University of Arizona by Academic Year",
       x = "Academic Year",
       y = "",
       color = "",
       caption = "Data from IPEDS Completions Survey") +
  theme_minimal() +
  theme(legend.position = "top")

ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total", unitid == 104179),
       mapping = aes(x = year,
                     y = n_deg,
                     color = factor(awlevelc,
                      levels = c("Bachelor's degree",
                                 "Master's degree",
                                 "Doctor's degree")))) +
  geom_line() +
  scale_y_continuous(labels = scales::comma) +
  scale_x_continuous(breaks = c(2017, 2018, 2019),
                     labels = c("2016-17", "2017-18", "2018-18")) +
  labs(title = "Number of Degrees Awarded by the University of Arizona by Academic Year",
       x = "Academic Year",
       y = "",
       color = "",
       caption = "Data from IPEDS Completions Survey") +
  theme_minimal() +
  theme(legend.position = "top")


# Add points on top of the lines and remove line legend
ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total", unitid == 104179),
       mapping = aes(x = year,
                     y = n_deg,
                     color = factor(awlevelc,
                      levels = c("Bachelor's degree",
                                 "Master's degree",
                                 "Doctor's degree")))) +
  geom_line(show.legend = FALSE) +
  geom_point() +
  scale_y_continuous(labels = scales::comma) +
  scale_x_continuous(breaks = c(2017, 2018, 2019),
                     labels = c("2016-17", "2017-18", "2018-18")) +
  labs(title = "Number of Degrees Awarded by the University of Arizona by Academic Year",
       x = "Academic Year",
       y = "",
       color = "",
       caption = "Data from IPEDS Completions Survey") +
  theme_minimal() +
  theme(legend.position = "top")

# Change color using ColorBrewer
ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total", unitid == 104179),
       mapping = aes(x = year,
                     y = n_deg,
                     color = factor(awlevelc,
                      levels = c("Bachelor's degree",
                                 "Master's degree",
                                 "Doctor's degree")))) +
  geom_line(show.legend = FALSE) +
  geom_point() +
  scale_y_continuous(labels = scales::comma) +
  scale_x_continuous(breaks = c(2017, 2018, 2019),
                     labels = c("2016-17", "2017-18", "2018-18")) +
  scale_color_brewer(palette = "Dark2") +
  labs(title = "Number of Degrees Awarded by the University of Arizona by Academic Year",
       x = "Academic Year",
       y = "",
       color = "",
       caption = "Data from IPEDS Completions Survey") +
  theme_minimal() +
  theme(legend.position = "top")

# Change color using manual colors
ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total", unitid == 104179),
       mapping = aes(x = year,
                     y = n_deg,
                     color = factor(awlevelc,
                      levels = c("Bachelor's degree",
                                 "Master's degree",
                                 "Doctor's degree")))) +
  geom_line(show.legend = FALSE) +
  geom_point() +
  scale_y_continuous(labels = scales::comma) +
  scale_x_continuous(breaks = c(2017, 2018, 2019),
                     labels = c("2016-17", "2017-18", "2018-18")) +
  scale_color_manual(values = c("#0C234B", "#AB0520", "#378DBD")) +
  labs(title = "Number of Degrees Awarded by the University of Arizona by Academic Year",
       x = "Academic Year",
       y = "",
       color = "",
       caption = "Data from IPEDS Completions Survey") +
  theme_minimal() +
  theme(legend.position = "top")





```

